### 2018-05-28 rigid_bodies_supp_info.tex
added supplementary material

\documentclass[
journal=jctcce,
manuscript=suppinfo]{achemso}

\setkeys{acs}{
	abbreviations = false,
	articletitle  = false,
	keywords      = false,
	maxauthors    = 10,
	super         = true
}

% Comment below before submitting
\let\titlefont\undefined
\usepackage[fontsize=11pt]{scrextend}
\usepackage[hidelinks,colorlinks,citecolor=blue]{hyperref}
%\flushbottom
% Comment above before submitting
\usepackage{booktabs}
\usepackage{threeparttable}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[T1]{fontenc}
\usepackage[table]{xcolor}
\definecolor{lightgray}{gray}{0.9}
\usepackage{tikz}
\usetikzlibrary{decorations.text}
\newcommand{\mt}[1]{\boldsymbol{\mathbf{#1}}}   % matrix symbol
\newcommand{\vt}[1]{\boldsymbol{\mathbf{#1}}}   % vector symbol
\newcommand{\tr}[1]{#1^\text{t}}                % transposition
\newcommand{\diff}[2]{\frac{\partial #1}{\partial #2}} % derivative
%\newcommand{\diff}[2]{\nabla_{#2}#1}            % derivative
\newcommand{\sdiff}[2]{\nabla^2_{#2}#1}         % 2nd derivative
\newcommand{\Ham}[1]{{\mathcal H}_\text{#1}}    % Hamiltonian
\newcommand{\Liu}[1]{i\!L_\text{#1}}            % Liouville
\newcommand{\timestep}{h}
\newcommand{\modified}[1]{\widetilde{#1}}

\usepackage{array}
\newcolumntype{L}{>{$}l<{$}}
\newcolumntype{C}{>{$}c<{$}}
\newcolumntype{R}{>{$}r<{$}}
\usepackage{booktabs}    
\usepackage[exponent-product=\times]{siunitx}

\newenvironment{smallarray}[1]                          % small arrays
{\null\,\vcenter\bgroup\scriptsize
	\renewcommand{\arraystretch}{1.5}
	\arraycolsep=.13885em
	\hbox\bgroup$\left[\array{@{}#1@{}}}
{\endarray\right]$\egroup\egroup\,\null}
\listfiles

\author{Ana J.
Silveira}
\email{asilveira@plapiqui.edu.ar}
\affiliation{Planta Piloto de Ingenier\'ia Qu\'imica, PLAPIQUI, Universidad Nacional del Sur, Camino La Carrindanga Km 7-CC: 717, Bah\'ia Blanca, Argentina}

\author{Charlles R.
A.
Abreu}
\email{abreu@eq.ufrj.br}
\affiliation{Chemical Engineering Department, Escola de Qu\'imica, Universidade Federal do Rio de Janeiro, Rio de Janeiro, RJ 21941-909, Brazil}

%\title{Revisiting the Nos\'{e}-Hoover chain thermostat for Rigid-Body Molecular Dynamics}
\title{Improvement of Deterministic NVT Molecular Dynamics via Backward Error Analysis}

%\abbreviations{AAA, BBB, CCC}
\keywords{aaa, bbb, ccc}

\begin{document}
\begin{suppinfo}

\begin{table}
	\caption{Effect of the time-step size on .... of 903 TIP3P\cite{Jorgensen_1983} water molecules in NVT MD simulations employing the unsplit solution for free rotations, given by the numerical schemes.}
    \label{table:stability_martyna}
    \begin{tabular}{SS[table-number-alignment = center,table-figures-exponent = 1]S[table-number-alignment = center,table-figures-exponent = 1]}
     & \multicolumn{2}{c}{Martyna \textit{et al.} \cite{Martyna_1996}}  \\
     \cmidrule[0.5mm](lr){2-3} 
     $\timestep$ $\text{(fs)}$  & $\modified{H}$  & $\modified{\Ham{}}$     \\
	1 &  -1.6e-2  & 4.5e-3\\
	2 & 9.5e-2 & 1.4e-1\\
	3 & 2.8e-1 & 5.6e-2\\
	4 & 1.1e0 & 5.6e-2\\
	5 & 3.3e0 & -3.7e-2\\
    6 & 6.1e1 & 6.9e-2 \\
	7 & 9.8e2 & -3.6e-3 \\
	8 & 1.1e4 & 1.8e-2 \\
	9 & 7.5e4 & 3.8e-3 \\
	10 & 3.3e5 & 1.9e-2  \\
	\end{tabular}
\end{table}

\end{suppinfo}

\bibliography{rigid_bodies}
\end{document}