### 2018-06-06 rigid_bodies_supp_info.tex

\documentclass[
journal=jctcce,
manuscript=suppinfo]{achemso}

\setkeys{acs}{
	abbreviations = false,
	articletitle  = false,
	keywords      = false,
	maxauthors    = 10,
	super         = true
}

% Comment below before submitting
\let\titlefont\undefined
\usepackage[fontsize=11pt]{scrextend}
\usepackage[hidelinks,colorlinks,citecolor=blue]{hyperref}
%\flushbottom
% Comment above before submitting
\usepackage{booktabs}
\usepackage{threeparttable}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[T1]{fontenc}
\usepackage[table]{xcolor}
\definecolor{lightgray}{gray}{0.9}
\usepackage{tikz}
\usetikzlibrary{decorations.text}
\newcommand{\mt}[1]{\boldsymbol{\mathbf{#1}}}   % matrix symbol
\newcommand{\vt}[1]{\boldsymbol{\mathbf{#1}}}   % vector symbol
\newcommand{\tr}[1]{#1^\text{t}}                % transposition
\newcommand{\diff}[2]{\frac{\partial #1}{\partial #2}} % derivative
%\newcommand{\diff}[2]{\nabla_{#2}#1}            % derivative
\newcommand{\sdiff}[2]{\nabla^2_{#2}#1}         % 2nd derivative
\newcommand{\Ham}[1]{{\mathcal H}_\text{#1}}    % Hamiltonian
\newcommand{\Liu}[1]{i\!L_\text{#1}}            % Liouville
\newcommand{\timestep}{h}
\newcommand{\modified}[1]{\widetilde{#1}}

\usepackage{array}
\newcolumntype{L}{>{$}l<{$}}
\newcolumntype{C}{>{$}c<{$}}
\newcolumntype{R}{>{$}r<{$}}
\usepackage{booktabs}    
\usepackage[exponent-product=\times]{siunitx}

\newenvironment{smallarray}[1]                          % small arrays
{\null\,\vcenter\bgroup\scriptsize
	\renewcommand{\arraystretch}{1.5}
	\arraycolsep=.13885em
	\hbox\bgroup$\left[\array{@{}#1@{}}}
{\endarray\right]$\egroup\egroup\,\null}
\listfiles

\author{Ana J.
Silveira}
\email{asilveira@plapiqui.edu.ar}
\affiliation{Planta Piloto de Ingenier\'ia Qu\'imica, PLAPIQUI, Universidad Nacional del Sur, Camino La Carrindanga Km 7-CC: 717, Bah\'ia Blanca, Argentina}

\author{Charlles R.
A.
Abreu}
\email{abreu@eq.ufrj.br}
\affiliation{Chemical Engineering Department, Escola de Qu\'imica, Universidade Federal do Rio de Janeiro, Rio de Janeiro, RJ 21941-909, Brazil}

%\title{Revisiting the Nos\'{e}-Hoover chain thermostat for Rigid-Body Molecular Dynamics}
\title{Improvement of Deterministic NVT Molecular Dynamics via Backward Error Analysis}

%\abbreviations{AAA, BBB, CCC}
\keywords{aaa, bbb, ccc}

\begin{document}
\begin{suppinfo}

\begin{table}
	\caption{Effect of the time-step size on .... of 903 TIP3P\cite{Jorgensen_1983} water molecules in NVT MD simulations employing the unsplit solution for free rotations, given by the numerical schemes.}
    \label{table:stability}
        \sisetup{table-number-alignment = center}
    \begin{tabular}{SS[table-figures-decimal=1,table-figures-exponent = 1]S[table-figures-exponent = 1,table-figures-decimal=1]S[table-figures-decimal=1,table-figures-exponent = 1]S[table-figures-exponent = 1,table-figures-decimal=1]S[table-figures-exponent = 1,table-figures-decimal=1]}
     & \multicolumn{5}{c}{Drift rate [(kcal/mol)/ns]}  \\
     & \multicolumn{2}{c}{This work} & \multicolumn{2}{c}{Martyna \textit{et al.} \cite{Martyna_1996}}   & \multicolumn{1}{c}{NVE \cite{Silveira_2017}} \\
     \cmidrule[0.5mm](lr){2-3} \cmidrule[0.5mm](lr){4-5} \cmidrule[0.5mm](lr){6-6}  
     $\timestep$ $\; \text{(fs)}$  & $\modified{H}$  & $\modified{\Ham{}}$  & $H$  & $\modified{\Ham{}}$     & $\modified{\Ham{}}$   \\
	1 &  4.6e-4  & 8.5e-2 & -1.6e-2  & 4.5e-3 & -1.e-3 \\
	2 & -8.2e-3 & -4.3e-2 &  9.5e-2 & 1.4e-1 &  9.3e-3 \\
	3 & -3.2e-2 & 7.3e-2 & 2.8e-1 & 5.6e-2 & -5.0e-3 \\
	4 & 8.1e-3 & -1.1e-1 & 1.1e0 & 5.6e-2 & 8.5e-2 \\
	5 & 1.3e0 & -2.2e-2 & 3.3e0 & -3.7e-2 & 1.1e0 \\
    6 & 4.4e1 & -3.4e-3 & 6.1e1 & 6.9e-2 & 6.0e1 \\
	7 & 8.8e2 & 2.2e-2  & 9.8e2 & -3.6e-3 & \text{-} \\
	8 & 9.8e3 & 7.6e-2  & 1.1e4 & 1.8e-2 & \text{-} \\
	9 & 7.1e4 & -2.5e-2  & 7.5e4 & 3.8e-3 & \text{-} \\
	10 & 3.2e5 & -3.2e-2  & 3.3e5 & 1.9e-2  & \text{-} \\
	\end{tabular}
\end{table}

ADD: COMPARE MILLER WITH AND WITHOUT RESPA
ADD: CORRELATION FUNCTION AS A FUNCTION OF MASS PARAMETER OF THERMOSTAT

\begin{table*}
	\caption{Effect of the time-step size on .... of 903 TIP3P\cite{Jorgensen_1983} water molecules in NVT MD simulations employing the unsplit solution for free rotations, given by the numerical schemes....}
	\label{table:check_ens}
	\sisetup{table-number-alignment = center}
    \begin{tabular}{SS[separate-uncertainty,table-figures-uncertainty = 2]SS[separate-uncertainty,table-figures-uncertainty= 1]SS[separate-uncertainty,table-figures-uncertainty= 1]S}
	{$\timestep$} & {estimated $\Delta T$} & {$\sigma$ deviations}  & {estimated $\Delta T$}  & {$\sigma$ deviations}  &    {estimated $\Delta T$}  & {$\sigma$ deviations} \\
	{(fs)} & {(K)} &   & {(K)}  &  & {(K)}  & \\
			\hline
			\multicolumn{7}{c}{This work (Unsplit)}\\
1.0	& 3.02(2) & 1.07 & & & &  \\
2.0 & 3.02(2) & 0.81 & & & &  \\
3.0 & 2.98(2) & 0.73 & & & &  \\
4.0 & 3.01(2) & 0.41 & & & &  \\
5.0 & 3.02(2) & 0.79 & & & &  \\
6.0 & 3.03(3) & 1.23 & & & &  \\
7.0 & 2.98(2) & 0.74 & & & &  \\
            \hline
			\multicolumn{7}{c}{Martyna \textit{et al.} \cite{Martyna_1996}}\\
1.0	& 2.98(2) & 0.66 & 3.01(2) & 0.49 & 2.96(3) & 0.90 \\
2.0 & 2.98(2) & 0.73 & 2.99(2) & 0.03 & 3.02(3) & 0.66 \\
3.0 & 2.95(2) & 2.33 & 3.01(2) & 0.65 & 2.91(3) & 2.47 \\
4.0 & 3.00(2) & 0.01 & 3.01(2) & 0.72 & 2.92(3) & 2.19  \\
5.0 & 2.93(2) & 3.05 & 3.01(2) & 0.70 & 2.94(3) & 1.58  \\
6.0 & 2.92(2) & 3.87 & 3.00(2) & 0.22 & 2.91(3) & 2.76  \\
7.0 & 2.89(2) & 5.35 & 3.02(1) & 1.33 & 2.87(2) & 4.33  \\
            \hline
			\multicolumn{7}{c}{Kamberaj \textit{et al.} \cite{Kamberaj_2005}}\\
1.0	& 2.97(2) & 1.16 & 2.99(2) & 0.16 & 2.89(3) & 2.84  \\
2.0 & 3.00(2) & 0.02 & 3.02(2) & 0.99 & 2.00(4) & 0.25  \\
3.0 & 2.99(4) & 0.18 & 3.08(4) & 4.09 & 3.1(1) & 1.17  \\
4.0 & 2.92(8) & 0.99 & 3.12(2) & 6.01 & 2.8(2) & 0.95  \\
5.0 & 3.02(5) & 0.40 & 3.2(2)  & 9.50 &  2.9(1) & 0.28  \\
6.0 & 3.14(5) & 2.69 & 3.29(2) & 14.68 & 3.21(7) & 2.90  \\
7.0 & 3.03(4) & 0.73 & 3.35(2) & 17.37 & 2.99(6) & 0.16  \\
    \end{tabular}
\end{table*}

\end{suppinfo}

\bibliography{rigid_bodies}
\end{document}